---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# staggeredGMM

<!-- badges: start -->
[![R-CMD-check](https://github.com/RishabhBijani/staggeredGMM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/RishabhBijani/staggeredGMM/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://app.codecov.io/gh/RishabhBijani/staggeredGMM/graph/badge.svg)](https://app.codecov.io/gh/RishabhBijani/staggeredGMM)
<!-- badges: end -->

`staggeredGMM` estimates cohort-by-time average treatment effects (CATTs)
under staggered treatment adoption by the generalized method of moments,
implementing the estimator of Arora and Bijani (2026). Only clean
two-by-two difference-in-differences comparisons -- against never-treated or
not-yet-treated controls -- enter the moment system.

## Installation

```r
# install.packages("devtools")
devtools::install_github("RishabhBijani/staggeredGMM")
```

## Usage

```{r example}
library(staggeredGMM)

fit <- gmm_staggered(sim_panel, yname = "y", tname = "year",
                     idname = "unit_id", gname = "cohort")
fit
```

```{r catt}
head(fit$catt)
```

## Weighting

The covariance model used to form the optimal weight is chosen with
`weighting`:

| `weighting` | Covariance model | Paper |
|---|---|---|
| `"pooled_toeplitz"` (default) | One stationary autocovariance sequence shared by every group | GMM-T |
| `"cohort_toeplitz"` | A separate stationary sequence per cohort | GMM-HT |
| `"unrestricted"` | Full unrestricted within-cohort covariance | GMM-U |

All three target the same CATT vector from the same moment conditions. They
do not generally return the same numbers: the system is over-identified, so
a different weighting matrix gives a different estimate.

## Testing the identifying assumptions

```{r jtest}
gmm_j_test(fit)
```

## Baseline covariates

`covar` applies the outcome-regression adjustment of Section 4.5 of the
paper, so that parallel trends need hold only conditional on the named
baseline covariates:

```{r covar}
fit_cov <- gmm_staggered(sim_panel, yname = "y", tname = "year",
                         idname = "unit_id", gname = "cohort",
                         covar = c("x1", "x2"))
fit_cov$aggregate$CW$estimate
```

## Bundled data

`sim_panel` is a simulated panel used in the examples and tests.
`beck_banks` is a real state-level panel from Beck, Levine and Levkov
(2010), included under CC BY 4.0; see `?beck_banks` and `inst/LICENSE.note`.
It contains thirteen always-treated states and is documented as a worked
example of that pitfall.

## Citation

```r
citation("staggeredGMM")
```

> Arora, P. and Bijani, R. (2026). "Estimating Treatment Effects under
> Staggered Timing and Non-Spherical Errors." Available at SSRN:
> <https://doi.org/10.2139/ssrn.6558759>

## License

MIT (c) Rishabh Bijani, Parush Arora. One bundled dataset carries its own
licence; see `inst/LICENSE.note`.
