---
title: "Getting started with staggeredGMM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with staggeredGMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(staggeredGMM)
```

## Overview

`staggeredGMM` estimates cohort x time-specific average treatment effects (CATTs) under staggered treatment adoption by generalized method of moments (GMM) as proposed by Arora and Bijani (2026).

> Arora, P. and Bijani, R. (2026). "Efficient Estimation of Treatment Effects under Staggered Adoption: GMM Approach." Ashoka University Economics Discussion Paper 163. <https://dp.ashoka.edu.in/ash/wpaper/paper163_0.pdf>

> Arora, P. and Bijani, R. (2026). "Estimating Treatment Effects under Staggered Timing and Non-Spherical Errors." Available at SSRN: <https://ssrn.com/abstract=6558759> or <http://dx.doi.org/10.2139/ssrn.6558759>.

Only clean comparisons (a treated cohort against a never-treated group, or against a not-yet-treated cohort) are used as moment conditions; already-treated units are never used as controls. The package provides three models:

- `gmm_staggered_I()`: pooled Toeplitz, one autocovariance sequence shared by every cohort.
- `gmm_staggered_II()`: cohort-specific Toeplitz, each cohort estimates its own sequence from its own residuals.
- `gmm_staggered_III()`: fully unrestricted within-cohort covariance, no stationarity assumed.

All three return the same cohort-and-time-specific CATTs and an aggregate ATT, differing only in how the weighting matrix that makes the GMM estimator efficient is constructed.

## The Example Data

`sim_panel` is a balanced synthetic panel bundled with the package: 60 units observed over 33 periods, with five treatment cohorts (first treated in periods 10, 13, 16, 19, or 22) and a never-treated group, cohort-heterogeneous dynamic treatment effects, AR(1)-correlated errors, and two baseline covariates, `x1` and `x2`.

```{r data}
str(sim_panel)
table(sim_panel$cohort[!duplicated(sim_panel$unit_id)])
```

## Running An Estimator

Every estimator takes the same core arguments: the outcome, time, unit, and cohort column names, and whether to include never-treated units as an additional control group.

```{r basic}
res <- gmm_staggered_I(sim_panel, y_var = "y", t_var = "year",
                        id_var = "unit_id", gname = "cohort",
                        has_nt = TRUE)
res$catt_out
res$ATT_CW
res$SE_CW
```

`catt_out` gives one row per (cohort, post-period) CATT. `ATT_CW` aggregates these into a single number, weighted by each cohort's share of treated observations; `SE_CW` is its standard error.

`eff_ok` records whether the iterated GMM reweighting converged:

```{r eff_ok}
res$eff_ok
res$n_iter
```

## Comparing The Three Estimators

```{r compare}
res2 <- gmm_staggered_II(sim_panel, y_var = "y", t_var = "year",
                          id_var = "unit_id", gname = "cohort", has_nt = TRUE)
res3 <- gmm_staggered_III(sim_panel, y_var = "y", t_var = "year",
                           id_var = "unit_id", gname = "cohort", has_nt = TRUE)
data.frame(estimator = c("I", "II", "III"),
           ATT_CW    = c(res$ATT_CW, res2$ATT_CW, res3$ATT_CW),
           SE_CW     = c(res$SE_CW, res2$SE_CW, res3$SE_CW),
           eff_ok    = c(res$eff_ok, res2$eff_ok, res3$eff_ok))
```

## Baseline-Covariate Adjustment

All three estimators accept an optional `covar` argument: a character vector naming baseline (pre-treatment) covariate columns. This applies the outcome-regression adjustment of Section 4.5 of the paper, which allows the parallel-trends assumption to hold only after conditioning on `covar` rather than unconditionally. More than one covariate can be supplied at once:

```{r covar}
res_cov <- gmm_staggered_I(sim_panel, y_var = "y", t_var = "year",
                           id_var = "unit_id", gname = "cohort",
                           has_nt = TRUE, covar = c("x1", "x2"))
res_cov$ATT_CW
res_cov$SE_CW
```

If a supplied covariate turns out not to be actually constant over time for some units (a baseline covariate is assumed to be pre-treatment and fixed, but `staggeredGMM` does not simply trust that), it uses each affected unit's first-observed value and raises a warning naming exactly which units were affected, rather than silently using a value that changes within a unit.

## Warnings And Errors

A handful of conditions raise a warning or an error rather than silently returning a misleading number; see `?gmm_staggered_I` (section "Warnings") for the full list.

## Other Bundled Datasets

Two real-world datasets ship with the package for reference: `beck_banks` (Beck, Levine, and Levkov 2010) and `cook_shah_nregs` (Cook and Shah 2022). See `?beck_banks` and `?cook_shah_nregs` for details.
