---
title: "Getting started with staggeredGMM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with staggeredGMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(staggeredGMM)
```

## Overview

`staggeredGMM` estimates cohort-by-time average treatment effects (CATTs)
under staggered treatment adoption by the generalized method of moments, as
proposed in

> Arora, P. and Bijani, R. (2026). "Estimating Treatment Effects under
> Staggered Timing and Non-Spherical Errors." Available at SSRN:
> <https://doi.org/10.2139/ssrn.6558759>

Only *clean* two-by-two difference-in-differences comparisons enter the
moment system: a treated cohort against a never-treated group, or against a
cohort not yet treated at the relevant period. Comparisons against
already-treated controls are exact linear combinations of the clean ones and
are never formed.

This vignette covers how to call the estimator and what it returns. For the
methodology, the choice among the three weighting schemes, and the
simulation and empirical evidence, see the paper.

## Data coding conventions

`gmm_staggered()` is strict about its input, and rejects rather than coerces.
The requirements are:

* one row per unit-period, with no duplicated unit-period pair;
* a numeric outcome (factor, character and logical outcomes are errors);
* integer-valued time and cohort columns;
* the cohort constant within unit, with `0` for never-treated units;
* no period absent from the whole panel;
* no unit treated at or before the first observed period.

Missing outcomes are allowed. Missing unit ids, periods or cohorts are not.

Two conventions are worth stating explicitly. First, a unit whose cohort
falls *after* the last observed period is untreated throughout the window,
so it is pooled with the never-treated group as a clean control rather than
discarded. Second, a unit treated at or before the first observed period has
no pre-treatment period at all, so none of its effects is identified; rather
than return zeros for those cells and contaminate the aggregate, the
estimator stops and names the offending cohort.

## A first estimate

`sim_panel` is a balanced synthetic panel: 60 units over 33 periods, five
treated cohorts and a never-treated group.

```{r basic}
fit <- gmm_staggered(sim_panel, yname = "y", tname = "year",
                     idname = "unit_id", gname = "cohort")
fit
```

The never-treated group was detected automatically. `never_treated = TRUE`
requires one and errors if none exists; `never_treated = FALSE` excludes
them.

## What is returned

```{r catt}
head(fit$catt)
```

`catt` gives one row per cohort-by-time cell, with the event time, the
estimate and its standard error, whether the cell is identified, and how
many clean comparisons anchor it. `coef()`, `vcov()` and `confint()` work as
usual:

```{r methods}
head(coef(fit))
head(confint(fit))
```

## The two aggregates

Two scalar summaries are reported, and they answer different questions.

* **`CW`, treated-observation weighting.** Every treated unit-period gets
  equal weight, so a cohort observed for longer contributes more total
  weight. This is the aggregation used throughout the paper, and it
  coincides with the "simple" aggregation of Callaway and Sant'Anna (2021).
* **`EW`, cohort-equal weighting.** Each treated cohort gets total weight
  `1/K`, spread equally over its own post-treatment window, so exposure
  length does not affect a cohort's influence.

```{r aggregates}
c(CW = fit$aggregate$CW$estimate, EW = fit$aggregate$EW$estimate)
```

On this design the true values are `-16.79375` and `-15.82020`
respectively; they differ because earlier cohorts are observed for more
post-treatment periods. Neither is more correct than the other. For any
other target, weight `fit$catt$estimate` directly.

## Choosing a weighting

```{r weighting}
fit_ht <- gmm_staggered(sim_panel, yname = "y", tname = "year",
                        idname = "unit_id", gname = "cohort",
                        weighting = "cohort_toeplitz")
c(pooled = fit$aggregate$CW$estimate, cohort = fit_ht$aggregate$CW$estimate)
```

All three weightings target the same CATT vector using the same moment
conditions. They do not generally return the same numbers: the system is
over-identified, so a different weighting matrix gives a different estimate.
Which to prefer is a methodological question the paper addresses.

## Convergence

```{r convergence}
unlist(fit$convergence[c("converged", "solve_ok", "termination", "n_iter")])
```

`converged` is `TRUE` only when the largest change in any CATT fell below
`tol`. It is deliberately distinct from `solve_ok`, which records merely that
a weighted solve succeeded; a run can solve successfully at every iteration
and still exhaust `max_iter` without converging. `termination` says why the
iteration stopped.

`max_iter` defaults to 100 because the iteration can converge linearly at a
slow rate. On this design `weighting = "cohort_toeplitz"` needs 86 steps,
though the aggregate ATT is stable to five significant figures well before
that — `tol` applies to the individual effects, which settle later.

The unrestricted weighting is a stronger constraint than it looks. Each
cohort's covariance is built from that cohort's residual vectors, so its rank
cannot exceed `min(N_g, T - 1)`, and the weight's total rank is bounded by
that quantity summed over groups. Unless the total comfortably exceeds the
number of effects, the normal equations are singular and no weighted step
completes at all. `sim_panel` has 60 units and 90 effects, so it does not qualify:

```{r unrestricted}
fit_u <- gmm_staggered(sim_panel, yname = "y", tname = "year",
                       idname = "unit_id", gname = "cohort",
                       weighting = "unrestricted")
```

The estimate returned in that case is the identity-weighted seed, not an
unrestricted-weighted GMM estimator, which is why it always warns. Failing to
*converge* after a weighted step did run is a different matter, and is common
enough under this weighting that it does not warn.

## Unbalanced panels

Missing outcomes are supported. Group-period means are taken over the units
observed at that period; autocovariances are divided by the number of
residual pairs actually observed at each lag; and the variance of a
group-period mean accounts for the overlap between the unit sets
contributing at each pair of periods. All three reduce exactly to the
complete-data formulas when nothing is missing, so there is one code path
rather than two.

```{r unbalanced}
gappy <- sim_panel
set.seed(1)
gappy$y[sample(nrow(gappy), 100)] <- NA
fit_gap <- gmm_staggered(gappy, yname = "y", tname = "year",
                         idname = "unit_id", gname = "cohort")
c(balanced = fit$aggregate$CW$estimate,
  unbalanced = fit_gap$aggregate$CW$estimate)
```

This is valid when missingness is independent of the outcome given group and
period. Selection on outcomes is not addressed by any weighting scheme.

A comparison referencing a group-period at which *no* unit is observed
cannot be formed and is dropped, with a warning giving the count.

## Partial identification

A cohort-by-time effect with no clean comparison is not estimable. Such
cells are reported as `NA` with `identified = FALSE`, never as an estimated
zero, and are excluded from the parameter vector so that one unestimable
cell does not disable the efficient weighting everywhere else.

Where an unidentified cell carries positive aggregation weight, the
aggregate itself is `NA`, and an identified-subset aggregate renormalised
over the estimable cells is reported alongside it:

```{r partial}
names(fit$aggregate$CW)
```

## Baseline covariates

`covar` takes a character vector of baseline, time-invariant covariate
columns and applies the outcome-regression adjustment of Section 4.5 of the
paper, so that parallel trends need hold only conditional on them.

```{r covar}
fit_cov <- gmm_staggered(sim_panel, yname = "y", tname = "year",
                         idname = "unit_id", gname = "cohort",
                         covar = c("x1", "x2"))
fit_cov$aggregate$CW$estimate
```

If a supplied covariate is not in fact constant within unit, each affected
unit's value at its **earliest observed period** is used and a warning names
the units. Earliest by time, not first by row: the answer does not depend on
how the data were sorted.

Where the adjustment cannot be identified for some control cohort and pair
of periods -- too few complete-case control units, or a rank-deficient fit
-- those comparisons alone revert to the unconditional form, and the count
is reported in `fit$controls$cov_n_fallback`.

## Testing the identifying assumptions

The over-identifying restrictions are the pre-treatment placebo moments,
whose expectation is zero only under parallel trends and no anticipation.
`gmm_j_test()` tests them jointly, weighting them by the estimated
serial-correlation model.

```{r jtest}
gmm_j_test(fit)
```

By default the test uses a local pre-window of three periods before each
cohort's adoption. Using every placebo restriction is also available, but a
large number of restrictions degrades the chi-squared approximation when the
number of units is small:

```{r jtest-full}
gmm_j_test(fit, type = "full")
```

Two cautions. Estimation and testing are deliberately decoupled -- the
estimate uses the full moment system, the test a local window -- because
using a pre-trends test to gate an estimate distorts its conditional
distribution (Roth, 2022). And the test speaks only to pre-treatment
periods: post-treatment parallel trends, the assumption that actually
identifies the effects, is untestable, so a clean result is necessary but
not sufficient.

The test requires a never-treated group, since the event-study basis it uses
differences every cohort against a common never-treated reference.

## A real design, and a coding pitfall

`beck_banks` is a panel of 49 US states over 1976--2006 on bank branch
deregulation and income inequality. It is also a worked example of the most
common way a staggered design goes wrong.

```{r beck-pitfall, error = TRUE}
gmm_staggered(beck_banks, yname = "ln_gini", tname = "wrkyr",
              idname = "state", gname = "branch_reform")
```

Thirteen states deregulated at or before 1976, the first observed year, ten
of them coded `1960`. They are treated throughout the window and have no
pre-treatment period, so none of their effects is identified. Silently
including them would return a plausible-looking aggregate contaminated by
cells that were never estimated. Restrict to states adopting inside the
window:

```{r beck-fixed}
dat <- beck_banks[beck_banks$branch_reform > 1976, ]
fit_beck <- gmm_staggered(dat, yname = "ln_gini", tname = "wrkyr",
                          idname = "state", gname = "branch_reform")
fit_beck
```

This design has no never-treated group either, so identification rests
entirely on not-yet-treated controls. The consequence is visible in the
output: the last cohort to adopt has no later cohort to serve as its
control, so none of its effects is identified, and neither is any cell at a
period after the final adoption year. Of the 380 cohort-by-time cells, 144
are unestimable.

```{r beck-identified}
table(fit_beck$catt$identified)
fit_beck$aggregate$CW$identified_weight_share
```

Because unidentified cells carry positive weight, the aggregate itself is
`NA` and the identified-subset figure is reported instead. That is the
intended behaviour: an aggregate over a set of cells more than a third of
which were never estimated would not mean what its name suggests. If a
never-treated group is available in your own application, supplying it
identifies far more of the surface.

## Citation

```{r citation, eval = FALSE}
citation("staggeredGMM")
```
